Manuscript

### 2015-04-15 inst/simulation/manuscript.Rmd
...

populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames[c(1,3,4,5)])
populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,4,5)])
populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,4,5)])

@@

meanPopulationSizes2000  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[c(1,3,5)]))
logPopulationSizes2000 <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[c(1,3,5)]))
logMeanPopulationSizes2000  <- getMeanPopulationSizes(logPopulationSizes2000)

meanPopulationSizes10Days  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[c(1,3,5)]))
logPopulationSizes10Days <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[c(1,3,5)]))
logMeanPopulationSizes10Days  <- getMeanPopulationSizes(logPopulationSizes10Days)
```

@@

    y <- lm(Estimated~Observed, x)
    prop <- x$Estimated/x$Observed
    prop <- prop[!is.infinite(prop) & prop > 0]
    prop.mean <- mean(log(prop))
    prop.sd <- sd(log(prop))
    prop.min <- quantile(log(prop), .025)
    prop.max <- quantile(log(prop), .975)
    slope.sd <- coef(summary(y))[2,2]

@@

table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=cutoff, models=modelNames[c(1,3,5)])
logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(cutoff+1), models=modelNames[1])
logTable1$Failed <- table1$Failed

table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=10*cutoff, models=modelNames[c(1,3,5)])
logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(10*cutoff+1), models=modelNames[1])
logTable2$Failed <- table2$Failed

table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=cutoff, models=modelNames[c(1,3,5)])
logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(cutoff+1), models=modelNames[1])
logTable3$Failed <- table3$Failed

@@

plotPopulationSizesTable <- function(logTable) {
  library(reshape2)
  facet_labeller <- function(variable, value) {
    measured <- list(Slope=expression(beta), R2=expression(R^2), EstTrue=expression(alpha), Failed="Failed (%)")
    print(paste0(variable,"=",value))
    if (variable=="variable") return(measured[value])
    if (variable=="Case") return(paste("Case", value))
  }
  x <- melt(logTable, id.vars=c("Scenario", "Model", "Case"), measure.vars=c("Slope", "R2", "EstTrue", "Failed"))
  bestFit <- data.frame(variable=unique(x$variable), y=c(1, 1, 0, 0))
  errorBars <- x
  errorBars$value <- NA
  #errorBars$value[errorBars$variable=="Slope"] <- logTable$Slope
  #errorBars$qmin[errorBars$variable=="Slope"] <- logTable$Slope-logTable$SlopeSD
  #errorBars$qmax[errorBars$variable=="Slope"] <- logTable$Slope+logTable$SlopeSD
  errorBars$value[errorBars$variable=="EstTrue"] <- logTable$EstTrue
  errorBars$qmin[errorBars$variable=="EstTrue"] <- logTable$EstTrueMin
  errorBars$qmax[errorBars$variable=="EstTrue"] <- logTable$EstTrueMax

  p <- ggplot(x, aes(Scenario, value, group=Model)) +
    geom_hline(data=bestFit, aes(yintercept=y), color="darkgray") + 
    geom_errorbar(data=errorBars, aes(Scenario, group=Model, ymax=qmax, ymin=qmin), position=position_dodge(width=.7), colour="darkgray") +
    geom_point(aes(shape=Model), position=position_dodge(width=.7)) +
    facet_grid(variable~Case, scales="free_y", labeller=facet_labeller) + ylab("") +
    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=10),
                                        legend.position="bottom",
                                        strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
                                        plot.margin=unit(c(0, 0, -.5, -.5), units="line"),
                                        legend.margin=unit(0, units="line"))# , legend.key=element_rect(fill="transparent"))
  return(p)
}

logTable1$Case <- "1"
logTable2$Case <- "2"
logTable3$Case <- "3"
logTable <- rbind(subset(logTable1, Model != "VT"), subset(logTable2, Model != "VT"), subset(logTable3, Model != "VT"))
p <- plotPopulationSizesTable(logTable)
saveFigure(p, "manuscript-figure-3.png", width=17, height=17, dpi=200) # draft quality
saveFigure(p, "manuscript-figure-3.pdf", width=17, height=17, dpi=300)

logTablePretty <- logTable
logTablePretty$Slope <- round(logTablePretty$Slope, 2)
logTablePretty$R2 <- round(logTablePretty$R2, 2)
logTablePretty$EstTrue <- round(logTablePretty$EstTrue, 2)
logTablePretty$EstTrueSD <- round(logTablePretty$EstTrueSD, 2)
logTablePretty$Failed <- round(logTablePretty$Failed, 0)
logTablePretty <- logTablePretty[,c("Case","Scenario","Model","Slope","R2","EstTrue","EstTrueSD","Failed")]
writeTable(subset(logTablePretty, Model != "VT"), "table-1.docx", context)

@@

@@

```{r figure-3sample}
xlim <- log10(c(30,800)); ylim <- xlim
logPopulationSizesSample <- subset(logPopulationSizes, scenario == "A")
logPopulationSizesSample$modelName <- prettyModelName(logPopulationSizesSample$modelName)
logPopulationSizesSample <- rbind(
  subset(logPopulationSizesSample, modelName=="ST"),
  subset(logPopulationSizesSample, modelName=="T1"),
  subset(logPopulationSizesSample, modelName=="T2"),
  subset(logPopulationSizesSample, modelName=="FMP"))

#p <- plotLogPopulationSize(subset(logPopulationSizesSample, modelName==modelNames[c(1,3,4)]), modelName=modelNames[c(1,3,4)], "a", xlim=xlim, ylim=ylim)

  #if (!any(abs(size$Estimated) == Inf)) p <- p + geom_smooth(method="lm", se=F, fullrange=T, colour="black")
  ls <- ddply(size, .(modelName), function(x) if (!any(abs(x$Estimated) == Inf)) coef(lm(Estimated~Observed, x)) else c(NA,NA))
  p <- p + geom_abline(data=ls, aes(intercept=`(Intercept)`, slope=Observed))

p <- plotLogPopulationSizeSample(subset(logPopulationSizesSample, modelName==c("ST","T1","T2","FMP")), modelName=c("ST","T1","T2","FMP"), "a", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3sample.png", height=7.87*2/3) # draft quality
saveFigure(p, "manuscript-figure-3sample.pdf", height=7.87*2/3, dpi=300)
```

@@

plotPopulationCorrelations <- function(populationCorrelations, case) {
  populationCorrelations$Scenario <- prettyScenario(populationCorrelations$Scenario)
  p <- ggplot(populationCorrelations, aes(True, Correlation)) + geom_point(size=1, colour="black") +
  facet_wrap(~Scenario, ncol=2) + xlab("True population size") +
  scale_y_continuous(limits=c(-1,1)) +
  ggtitle(paste("Case:", case)) +
  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
                                      plot.title=element_text(size=10),
                                      plot.margin=unit(c(0.2, 0, 0, 0), units="line"),
                                      panel.margin=unit(0, units="line"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
  return(p)
}

@@

plotCIProportions <- function(credibilityIntervals) {
  modelLabels <- prettyModelName(unique(credibilityIntervals$modelName))
  credibilityIntervals$CI <- credibilityIntervals$Interval
  p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=CI, colour=modelName)) +
    geom_hline(aes(yintercept=.5), color="grey90") +
    geom_hline(aes(yintercept=.95), color="grey90") +
    geom_point() +
    scale_shape_manual(values=c(16, 1)) +
    scale_colour_manual("Model", values=c("black", "gray26", "gray80"), labels=modelLabels) +
    #scale_colour_manual("Model", values=c("blue", "green", "red"), labels=modelLabels) +
    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
    xlab("Scenario") + ylab("Proportion") +
    facet_wrap(~panel) + ylim(0, 1) +
    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
          plot.margin=unit(c(0.2, 0, 0, 0), units="line"),
          #legend.background=element_rect(fill="transparent"),
          legend.margin=unit(-2, units="line"),
          panel.margin=unit(0, units="line"),
          legend.position="bottom",
          legend.box="horizontal",
          legend.key.width=unit(.7, "line"))
  return(p)
}
#p <- plotCIProportions(cip)
#saveFigure(p, "manuscript-figure-5.png", height=7.87/2, dpi=200)

cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[c(1,3,5)], panel="Case 1", scenarioNames=scenarios, cutoff=cutoff)
cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,5)], panel="Case 2", scenarioNames=scenarios, cutoff=10*cutoff)
cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,5)], panel="Case 3", scenarioNames=scenarios, cutoff=cutoff)

cip <- subset(rbind(cip1, cip2, cip3), modelName != "SmoothModel-nbinomial-ar1")
#cip <- subset(rbind(cip1, cip2, cip3), !(modelName %in% c("SmoothModel-nbinomial-ar1", "SmoothModel-nbinomial-matern-ar1")))
p <- plotCIProportions(cip)
saveFigure(p, "manuscript-figure-5.png", height=7.87/2)
saveFigure(p, "manuscript-figure-5.pdf", height=7.87/2, dpi=300)
```

>>>>

populationSizes <- getPopulationSizes(scenarios=scenarios, modelNames=modelNames[c(1,3,4,5)])
#populationSizes2000 <- getPopulationSizes(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,4,5)])
#populationSizes10Days <- getPopulationSizes(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,4,5)])

@@

#meanPopulationSizes2000  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[c(1,3,5)]))
#logPopulationSizes2000 <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes2000, 10*cutoff, models=modelNames[c(1,3,5)]))
#logMeanPopulationSizes2000  <- getMeanPopulationSizes(logPopulationSizes2000)

#meanPopulationSizes10Days  <- getMeanPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[c(1,3,5)]))
#logPopulationSizes10Days <- getLogPopulationSizes(cutoffPopulationSizes(populationSizes10Days, cutoff, models=modelNames[c(1,3,5)]))
#logMeanPopulationSizes10Days  <- getMeanPopulationSizes(logPopulationSizes10Days)
```

@@

    y <- lm(Estimated~Observed, x)
    #prop <- x$Estimated/x$Observed
    prop <- (x$Estimated+1)/(x$Observed+1)
    prop <- prop[!is.infinite(prop) & prop > 0]
    prop.mean <- mean(log(prop))
    prop.sd <- sd(log(prop))
    prop.min <- quantile(log(prop), .025)
    prop.max <- quantile(log(prop), .975)
    slope.sd <- coef(summary(y))[2,2]

@@

table1 <- getPopulationSizesTable(populationSizes, meanPopulationSizes, overEstimate=cutoff, models=modelNames[c(1,3,5)])
logTable1 <- getPopulationSizesTable(logPopulationSizes, logMeanPopulationSizes, overEstimate=log10(cutoff+1), models=modelNames[c(1,3,5)])
logTable1$Failed <- table1$Failed

#table2 <- getPopulationSizesTable(populationSizes2000, meanPopulationSizes2000, overEstimate=10*cutoff, models=modelNames[c(1,3,5)])
#logTable2 <- getPopulationSizesTable(logPopulationSizes2000, logMeanPopulationSizes2000, overEstimate=log10(10*cutoff+1), models=modelNames[1])
#logTable2$Failed <- table2$Failed

#table3 <- getPopulationSizesTable(populationSizes10Days, meanPopulationSizes10Days, overEstimate=cutoff, models=modelNames[c(1,3,5)])
#logTable3 <- getPopulationSizesTable(logPopulationSizes10Days, logMeanPopulationSizes10Days, overEstimate=log10(cutoff+1), models=modelNames[1])
#logTable3$Failed <- table3$Failed

@@

facet_wrap_labeller <- function(gg.plot,labels=NULL) {
  #works with R 3.0.1 and ggplot2 0.9.3.1
  require(gridExtra)

  g <- ggplotGrob(gg.plot)
  gg <- g$grobs      
  strips <- grep("strip_t", names(gg))

  for(ii in seq_along(labels))  {
    modgrob <- getGrob(gg[[strips[ii]]], "strip.text", 
                       grep=TRUE, global=TRUE)
    gg[[strips[ii]]]$children[[modgrob$name]] <- editGrob(modgrob,label=labels[ii])
  }

  g$grobs <- gg
  class(g) = c("arrange", "ggplot",class(g)) 
  g
}

plotPopulationSizesTable <- function(logTable) {
  library(reshape2)
  facet_labeller <- function(variable, value) {
    measured <- list(EstTrue=expression(alpha), Slope=expression(beta), R2=expression(R^2), Failed="Failed (%)")
    print(paste0(variable,"=",value))
    if (variable=="variable") return(measured[value])
    if (variable=="Case") return(paste("Case", value))
  }
  #x <- melt(logTable, id.vars=c("Scenario", "Model", "Case"), measure.vars=c("EstTrue", "Slope", "R2", "Failed"))
  x <- melt(logTable, id.vars=c("Scenario", "Model"), measure.vars=c("EstTrue", "Slope", "R2", "Failed"))

  bestFit <- data.frame(variable=unique(x$variable), y=c(0, 1, 1, 0))
  errorBars <- x
  errorBars$value <- NA
  #errorBars$value[errorBars$variable=="Slope"] <- logTable$Slope
  #errorBars$qmin[errorBars$variable=="Slope"] <- logTable$Slope-logTable$SlopeSD
  #errorBars$qmax[errorBars$variable=="Slope"] <- logTable$Slope+logTable$SlopeSD
  errorBars$value[errorBars$variable=="EstTrue"] <- logTable$EstTrue
  errorBars$qmin[errorBars$variable=="EstTrue"] <- logTable$EstTrueMin
  errorBars$qmax[errorBars$variable=="EstTrue"] <- logTable$EstTrueMax

  p <- ggplot(x, aes(Scenario, value, group=Model)) +
    geom_hline(data=bestFit, aes(yintercept=y), color="darkgray") + 
    geom_errorbar(data=errorBars, aes(Scenario, group=Model, ymax=qmax, ymin=qmin), position=position_dodge(width=.7), colour="darkgray") +
    geom_point(aes(shape=Model), position=position_dodge(width=.7)) + ylab("") +
    #facet_grid(variable~., scales="free_y", labeller=facet_labeller) +
    facet_wrap(~variable, scales="free_y", nrow=2) +
    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"), legend.title=element_text(size=10),
                                        legend.position="bottom",
                                        strip.text.x=element_text(size=10), strip.text.y=element_text(size=10),
                                        plot.margin=unit(c(0, 0.2, 0, -.5), units="line"),  #top, right, bottom, and left margins
                                        legend.margin=unit(-1.5, units="line"))# , legend.key=element_rect(fill="transparent"))

  # Expand y-axis 100 for Failed
  x <- data.frame(Scenario="A", Model="FMP", variable="Failed", value=100)
  p <- p + geom_blank(data=x, aes(x=Scenario, y=value, group=Model))

  p <- facet_wrap_labeller(p, labels=c(expression(alpha), Slope=expression(beta), R2=expression(R^2), Failed="Failed (%)"))

#logTable1$Case <- "1"
#logTable2$Case <- "2"
#logTable3$Case <- "3"
#logTable <- rbind(subset(logTable1, Model != "VT"), subset(logTable2, Model != "VT"), subset(logTable3, Model != "VT"))
logTable <- logTable1
p <- plotPopulationSizesTable(logTable)
saveFigure(p, "manuscript-figure-3.png", width=17, dpi=200) # draft quality
saveFigure(p, "manuscript-figure-3.pdf", height=17, dpi=300)

logTablePretty <- logTable
logTablePretty$Slope <- round(logTablePretty$Slope, 2)
logTablePretty$R2 <- round(logTablePretty$R2, 2)
logTablePretty$EstTrue <- round(logTablePretty$EstTrue, 2)
logTablePretty$EstTrueSD <- round(logTablePretty$EstTrueSD, 2)
logTablePretty$Failed <- round(logTablePretty$Failed, 0)
#logTablePretty <- logTablePretty[,c("Case","Scenario","Model","Slope","R2","EstTrue","EstTrueSD","Failed")]
logTablePretty <- logTablePretty[,c("Scenario","Model","Slope","R2","EstTrue","EstTrueSD","Failed")]
writeTable(subset(logTablePretty, Model != "VT"), "table-1.docx", context)

@@

```{r figure-3sample}
xlim <- log10(c(30,800)); ylim <- xlim
logPopulationSizesSample <- subset(logPopulationSizes, scenario == "A")
logPopulationSizesSample <- cutoffPopulationSizes(logPopulationSizesSample, cutoff, models=modelNames[c(1,3,5)])
logPopulationSizesSample$modelName <- prettyModelName(logPopulationSizesSample$modelName)
logPopulationSizesSample <- rbind(
  subset(logPopulationSizesSample, modelName=="ST"),
  subset(logPopulationSizesSample, modelName=="T1"),
  subset(logPopulationSizesSample, modelName=="T2"),
  subset(logPopulationSizesSample, modelName=="FMP"))

lm(Estimated~Observed, subset(logPopulationSizes, modelName=="FMPModel" & Scenario=="A"))
lm(Estimated~Observed, subset(logPopulationSizesSample, modelName=="FMP" & Scenario=="A"))

  #if (!any(abs(size$Estimated) == Inf)) p <- p + geom_smooth(method="lm", se=F, fullrange=T, colour="black")
  ls <- ddply(size, .(modelName), function(x) if (!any(abs(x$Estimated) == Inf)) coef(lm(Estimated~Observed, x)) else c(NA,NA))
  print(ls)
  p <- p + geom_abline(data=ls, aes(intercept=`(Intercept)`, slope=Observed))

p <- plotLogPopulationSizeSample(subset(logPopulationSizesSample, modelName %in% c("ST","T1","T2","FMP")), modelName=c("ST","T1","T2","FMP"), "a", xlim=xlim, ylim=ylim)
saveFigure(p, "manuscript-figure-3sample.png", height=7.87) # draft quality
saveFigure(p, "manuscript-figure-3sample.pdf", height=7.87, dpi=300)
```

@@

@@

plotPopulationCorrelations <- function(populationCorrelations, case) {
  populationCorrelations$Scenario <- prettyScenario(populationCorrelations$Scenario)
  p <- ggplot(populationCorrelations, aes(True, Correlation)) + geom_point(size=1, colour="black") +
  facet_wrap(~Scenario, ncol=2) + xlab("True population size") +
  scale_y_continuous(limits=c(-1,1)) +
  #ggtitle(paste("Case:", case)) +
  theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent"),
                                      plot.title=element_text(size=10),
                                      plot.margin=unit(c(0, 0.2, 0, 0), units="line"),
                                      panel.margin=unit(0, units="line"))#, axis.text.x=element_text(angle=45, hjust=.5, vjust=1))
  return(p)
}

@@

cip1 <- getCIProportions(scenarios=scenarios, modelNames=modelNames[c(1,3,5)], panel="Case 1", scenarioNames=scenarios, cutoff=cutoff)
#cip2 <- getCIProportions(scenarios=paste0(scenarios, "combined"), modelNames=modelNames[c(1,3,5)], panel="Case 2", scenarioNames=scenarios, cutoff=10*cutoff)
#cip3 <- getCIProportions(scenarios=paste0(scenarios, "10days"), modelNames=modelNames[c(1,3,5)], panel="Case 3", scenarioNames=scenarios, cutoff=cutoff)

plotCIProportions <- function(credibilityIntervals) {
  #modelLabels <- prettyModelName(unique(credibilityIntervals$modelName))
  credibilityIntervals$Model <- prettyModelName(credibilityIntervals$modelName)
  credibilityIntervals$CI <- credibilityIntervals$Interval
  p <- ggplot(credibilityIntervals, aes(scenario, Proportion, shape=Model, colour=CI)) +
    geom_hline(aes(yintercept=.5), color="grey90") +
    geom_hline(aes(yintercept=.95), color="grey90") +
    geom_point(aes(shape=Model, group=Model), position=position_dodge(width=.7)) +
    scale_shape_manual("Model", values=c(17,15,3)) +
    scale_colour_manual("CI", values=c("gray", "black")) +
    #scale_colour_manual("Model", values=c("blue", "green", "red"), labels=modelLabels) +
    theme_minimal(base_size=10) + theme(panel.border=element_rect(fill="transparent")) +
    xlab("Scenario") + ylab("Proportion") + ylim(0, 1) +
    #facet_wrap(~panel) +
    theme(panel.grid.major=element_blank(), panel.grid.minor=element_blank(),
          plot.margin=unit(c(0.2, 0.2, 0, 0), units="line"),
          #legend.background=element_rect(fill="transparent"),
          legend.margin=unit(-2, units="line"),
          panel.margin=unit(0, units="line"),
          legend.position="bottom",
          legend.box="horizontal",
          legend.key.width=unit(.7, "line"))
  return(p)
}

cip <- cip1
#cip <- subset(rbind(cip1, cip2, cip3), modelName != "SmoothModel-nbinomial-ar1")
#cip <- subset(rbind(cip1, cip2, cip3), !(modelName %in% c("SmoothModel-nbinomial-ar1", "SmoothModel-nbinomial-matern-ar1")))
p <- plotCIProportions(cip)
saveFigure(p, "manuscript-figure-5.png", height=7.87/2)
saveFigure(p, "manuscript-figure-5.pdf", height=7.87/2, dpi=300)
```
